A Flux Scale for Southern Hemisphere 21cm EoR Experiments

### 2013-04-23 ms.tex
Significant rewrite of S1.  S2 is a complete hodgepodge for now.

Numerous radio telescopes are now exploring the prospects for using
measurements of highly redshifted 21cm emission to inform our understanding of
cosmic reionization in the redshift range $6< z<12$, corresponding to radio
frequencies below 200 MHz (see reviews in
\citealt{Furlanetto:2006p2267,Morales:2010p8093,Pritchard:2012p9555}) .  These
include telescopes aiming to measure the global temperature change of 21cm
emission during the Epoch of Reionization (EoR), such as the Compact
Reionization Experiment (CoRE) and the Experiment to Detect the Global EoR
Signature (EDGES; \citealt{Bowman:2010p8546}), and interferometers aiming to
measure the power spectrum of 21-cm EoR emission, such as the Giant Metre-wave
Radio Telescope (GMRT;
\citealt{Paciga:2011p9470,Paciga:2013p9627})\footnote{\url{http://gmrt.ncra.tifr.res.in/}},
the LOw Frequency ARray (LOFAR;
\citealt{Yatawatta:2013p9699})\footnote{\url{http://www.lofar.org/}}, the
Murchison Widefield Array (MWA;  \citealt{Bowman:2012p9138};
\citealt{Tingay:2013p9022}))\footnote{\url{http://www.mwatelescope.org/}}, and
the Donald C. Backer Precision Array for Probing the Epoch of Reionization
(PAPER; \citealt{Parsons:2010p6757,Parsons2013b})\footnote{\url{http://eor.berkeley.edu/}}.

As such, there has been a renewed interest in the spectral and spatial variation of
foreground emission in the 100--200 MHz frequency band that covers
the $z=6$--13 redshift range expected to encompass reionization
\citep{furlanetto_et_al2006}.
In particular, the spectral properties of extra-galactic point-sources are important
both because they are valuable calibration references
and because they are strong foreground emitters that must be removed from 21-cm EoR measurements.
With the sparse availability of measured foreground properties in the
100--200-MHz frequency band over large areas of the sky \citep{deolivieracosta_et_al2008}, continued
foreground characterization is a
vital step en route to any 21-cm EoR detection.
Both PAPER and the MWA are located in the southern hemisphere at radio quiet
reserves being prepared for the upcoming Square Kilometer Array.  At these low
frequencies the Southern sky is much less well known than the North, with the
most extensive surveying work being conducted  by the EoR experiments
themselves \citep{Jacobs:2011p8438,Williams:2012p8768}.

% XXX is the "eor science" part necessary?  this isn't a paper on eor
Neutral hydrogen is visible during reionization as $\sim$30mK excess surface
brightness over the CMB punctuated both spectrally and spatially by regions of
zero emission.  This patchy signal is predicted to be visible everywhere on the
% XXX ARP: don't like description of EoR signal as "surface brightness"
sky at size scales up to degrees, or about $0.1$cMpc$^{-1}$.   To increase the
surface brightness sensitivity without building costly dishes, MWA, PAPER and
LOFAR have chosen to use combinations of phased and/or correlated dipoles. By
design also cover much wider fields of view ($>10$\arcdeg) and bandwidths
($\sim 100$\% fractional) than traditional dish telescopes.   As the elements
% XXX what about GMRT?
can not be physically pointed, the apparent flux scale varies across the sky.
This direction dependent gain is currently uncertain to 20\% or higher
\citep{Jacobs:2013p9713}.  The ideal solution is to use a wide-grid of
well-known calibrators to map the primary beam and set the flux scale.  
% XXX don't go to "ideal solution" just yet
% XXX ARP: add importance of accurate flux vs. frequency (Parsons et al. 2012, etc.)

@@

However, catalog source fluxes at 150MHz are also inaccurate to the 20\% level,
particularly at southern declinations ($\delta<-20$\arcdeg) . Historically, the best
EoR band data were by \citet{Slee:1995p7541} with Culgoora Circular
Array\footnote{Known during daylight hours as Culgoora Radio Heliograph} and
various higher frequency measurements with Parkes.  These data are typically
uncertain to 10\% or higher and provide little coverage of the EoR band beyond
a single narrow-band data point.

@@

The data  include new measurements of Pictor A that are uncertain to 3.5\%,
close to the level of the most precise higher frequency measurements by
\citet{Wills:1975p9314}. Using these new measurements we update the spectral
calibration model to an accuracy of 2.5\%. As validation of this method, we
also measure a selection of 58 additional sources at the same Declination.  The
primary beam is calibrated using a single common source (J2331-416), thus
limiting the declination probed to $-45 \pm 5$\arcdeg.  The flux scale is set
by Monte Carlo fitting to a subset of sources, all of which are calibrated to
the Baars et al scale. In \S\ref{sec:Observations} we describe the instrumental
setup, observations and analysis method followed. In \S\ref{sec:fits} we
compare the spectra with available data followed by conclusions in
\S\ref{sec:Conclusion}.

%This declination limit removes the necessity of correcting for a primary beam model and allows us to calculate more precise fluxes that provide a reference standard for future work.

>>>>

Numerous radio telescopes are now exploring the prospects for using
measurements of highly redshifted 21cm emission to inform our understanding of
cosmic reionization in the redshift range $6< z<12$, corresponding to radio
frequencies below 200 MHz (see reviews in
\citealt{Furlanetto:2006p2267,Morales:2010p8093,Pritchard:2012p9555}).  These
include telescopes aiming to measure the global temperature change of 21cm
emission during the Epoch of Reionization (EoR), such as the Compact
Reionization Experiment (CoRE) and the Experiment to Detect the Global EoR
Signature (EDGES; \citealt{Bowman:2010p8546}), and interferometers aiming to
measure the power spectrum of 21-cm EoR emission, such as the Giant Metre-wave
Radio Telescope (GMRT;
\citealt{Paciga:2011p9470,Paciga:2013p9627})\footnote{\url{http://gmrt.ncra.tifr.res.in/}},
the LOw Frequency ARray (LOFAR;
\citealt{Yatawatta:2013p9699})\footnote{\url{http://www.lofar.org/}}, the
Murchison Widefield Array (MWA;  \citealt{Bowman:2012p9138};
\citealt{Tingay:2013p9022}))\footnote{\url{http://www.mwatelescope.org/}}, and
the Donald C. Backer Precision Array for Probing the Epoch of Reionization
(PAPER; \citealt{Parsons:2010p6757,Parsons2013b})\footnote{\url{http://eor.berkeley.edu/}}.

As such, there has been a renewed interest in the spectral and spatial variation of
foreground emission in the 100--200 MHz frequency band that covers
the $z=6$--13 redshift range expected to encompass reionization
\citep{furlanetto_et_al2006}.
In particular, the spectral properties of extra-galactic point-sources are important
both because they are valuable calibration references
and because they are strong foreground emitters that must be removed from 21-cm EoR measurements.
With the sparse availability of measured foreground properties in the
100--200-MHz frequency band over large areas of the sky \citep{deolivieracosta_et_al2008}, continued
foreground characterization is a
vital step en route to any 21-cm EoR detection.
At these low
frequencies the Southern sky is much less well known than the North, 
with catalog source fluxes at 150MHz being inaccurate at the 20\% level
at southern declinations ($\delta<-20$\arcdeg) (XXX cite).
Both PAPER and the MWA are located in the southern hemisphere at radio-quiet
reserves being prepared for the upcoming Square Kilometer Array, and hence,
the
most extensive surveying work is now being conducted by the EoR experiments
themselves \citep{Jacobs:2011p8438,Williams:2012p8768}.

One significant complication to improving the state of affairs in foreground characterization
is that many 21cm EoR experiments, including LOFAR in the northern hemisphere, and
PAPER and MWA in the southern hemisphere, are designed for drift-scan observations.
This design decision has largely been driven by the simplicity and cost-effectiveness of
phased and/or correlated dipoles for achieving the aggressive sensitivity requirements
for measuring the 21cm power spectrum of reionization (XXX cite).
Adding to the challenge, these telescopes 
cover much wider fields of view ($>10$\arcdeg) and bandwidths
($\sim 100$\% fractional) than traditional dish telescopes.
Because they do not physically point, flux calibration for such arrays relies heavily
on an accurate model of the primary beam response to correct for an
the apparent flux scale that varies across the sky.
This direction dependent gain is currently uncertain to 20\% or higher
% XXX 20% seems a bit grim.  Doesn't jonnie argue for 5 to 10% in his paper?
\citep{Jacobs:2013p9713}, approximately matching the level of uncertainty in the catalogs.

In this paper, we set out to significantly improve the accuracy of spectral measurements between 
100--200 MHz for a set of bright sources in the declination range XXX to XXX that
are of particular value for southern-hemisphere 21cm EoR experiments such as PAPER and the MWA.
Using the fact that, for this restricted declination range,
sources transit through a nearly identical primary beam response pattern, we are able to avoid one of the
most debilitating source of error in these measurements: the primary beam.  
% XXX document map
In \S\ref{sec:analysis} we describe our approach for measuring source spectra with
drift-scan observations and deriving an absolute flux scale from catalog data.  
In \S\ref{sec:Observations} we detail the instrumental
setup, observations and analysis method followed. In \S\ref{sec:fits} we
compare the spectra with available data.  We conclude in 
\S\ref{sec:Conclusion}.

%The ideal solution is to use a wide-grid of
%well-known calibrators to map the primary beam and set the flux scale.  
% XXX don't go to "ideal solution" just yet
% XXX ARP: add importance of accurate flux vs. frequency (Parsons et al. 2012, etc.)

@@

However, catalog source fluxes at 150MHz are also inaccurate to the 20\% level,
particularly at southern declinations ($\delta<-20$\arcdeg).

Historically, the best
EoR band data were by \citet{Slee:1995p7541} with Culgoora Circular
Array\footnote{Known during daylight hours as Culgoora Radio Heliograph} and
various higher frequency measurements with Parkes.  These data are typically
uncertain to 10\% or higher and provide little coverage of the EoR band beyond
a single narrow-band data point.

@@

The data  include new measurements of Pictor A that are uncertain to 3.5\%,
close to the level of the most precise higher frequency measurements by
\citet{Wills:1975p9314}. Using these new measurements we update the spectral
calibration model to an accuracy of 2.5\%. As validation of this method, we
also measure a selection of 58 additional sources at the same Declination.  The
primary beam is calibrated using a single common source (J2331-416), thus
limiting the declination probed to $-45 \pm 5$\arcdeg.  The flux scale is set
by Monte Carlo fitting to a subset of sources, all of which are calibrated to
the Baars et al scale. 
%This declination limit removes the necessity of correcting for a primary beam model and allows us to calculate more precise fluxes that provide a reference standard for future work.
